### 2024-12-09 manuscript.tex
Addressing comments

% Related publications:
%
% Takasuka et al. (2024), A protocol and analysis of year-long simulations of global storm-resolving models and beyond
% Knight et al. (2024), Cloud properties and boundary layer stability above Southern Ocean sea ice and coastal Antarctica
% Zheng et al. (2024), Using Satellite and ARM Observations to Evaluate Cold Air Outbreak Cloud Transitions in E3SM Global Storm-Resolving Simulations
% Schmidt et al. (2024), Effects of vertical grid spacing on the climate simulated in the ICON-Sapphire global storm-resolving model
% Fons et al. (2024), Investigating the sign of stratocumulus adjustments to aerosols in the global storm-resolving model ICON
% Fiddes et al. (2024), A machine learning approach for evaluating Southern Ocean cloud radiative biases in a global atmosphere model
% Mooers et al. (2023), Comparing storm resolving models and climates via unsupervised machine learning
% Zhang et al. (2023), Understanding models' global sea surface temperature bias in mean state: from CMIP5 to CMIP6
% Grundner (2023), Data-Driven Cloud Cover Parameterizations for the ICON Earth System Model Using Deep Learning and Symbolic Regression [thesis]
% Luo et al. (2023), Origins of Southern Ocean warm sea surface temperature bias in CMIP6 models
% Sherriff-Tadano et al. (2023), Southern Ocean Surface Temperatures and Cloud Biases in Climate Models Connected to the Representation of Glacial Deep Ocean Circulation
% Danker et al. (2022), Exploring relations between cloud morphology, cloud phase, and cloud radiative properties in Southern Ocean's stratocumulus clouds
% Possner et al. (2022), Resolution Dependence of Southern Ocean Mixed-Phase Clouds in ICON
% Mauritsen et al. (2022), Early Development and Tuning of a Global Coupled Cloud Resolving Model, and its Fast Response to Increasing CO2
% Wall et al. (2022), Observational Constraints on Southern Ocean Cloud-Phase Feedback
% Grundner et al. (2022), Deep learning based cloud cover parameterization for ICON
% Cesana et al. (2022), Southern Ocean Solar Reflection Biases in CMIP6 Models Linked to Cloud Phase and Vertical Structure Representations
% Fiddes et al. (2022), Southern Ocean cloud and shortwave radiation biases in a nudged climate model simulation: does the model ever get it right?
% Zhao et al. (2022), Compensating Errors in Cloud Radiative and Physical Properties over the Southern Ocean in the CMIP6 Climate Models
% Konsta et al. (2022), Low-Level Marine Tropical Clouds in Six CMIP6 Models Are Too Few, Too Bright but Also Too Compact and Too Homogeneous
% Lin and Yu (2022), The potential impact of model horizontal resolution on the simulation of atmospheric cloud radiative effect in CMIP6 models
% Ramadoss et al. (2021), An evaluation of kilometre scale ICON simulations of mixed-phase stratocumulus over the Southern Ocean durin CAPRICORN [poster]
% Miao et al. (2021), A Regime-based Investigation into the Errors of CMIP6 Simulated Cloud Radiative Effects Using Satellite Observations
% Roh et al. (2021), Intercomparison of cloud properties in DYAMOND simulations over the Atlantic Ocean
% Caldwell et al. (2021), Convection-Permitting Simulations With the E3SM Global Atmosphere Model
% Schuddeboom and McDonald (2021), The Southern Ocean Radiative Bias, Cloud Compensating Errors, and Equilibrium Climate Sensitivity in CMIP6 Models
% Ramadoss et al. (2020), Simulating mixed-phase cloud properties with ICON around the CAPRICORN field campaign at the kilometre scale [poster]
% Heim et al. (2021), Inter-model Variability in Convection-Resolving Simulations of Subtropical Marine Low Clouds
% Seiki and Roh (2020), Improvements in Supercooled Liquid Water Simulations of Low-Level Mixed-Phase Clouds over the Southern Ocean Using a Single-Column Model
% Stevens et al. (2020), The added value of large-eddy and storm-resolving models for simulating clouds and precipitation
% Atlas et al. (2020), How well do large‐eddy simulations and global climate models represent observed boundary layer structures and low clouds over the summertime Southern Ocean?
% Korn et al. (2020), ICON‐O: The Ocean Component of the ICON Earth System Model—Global Simulation Characteristics and Local Telescoping Capability
% Vignesh et al. (2020), Assessment of CMIP6 Cloud Fraction and Comparison with Satellite Observations
% Chen et al. (2022), Evaluation of Simulated Cloud Diurnal Variation in CMIP6 Climate Models
% Jian et al. (2020), Evaluation of the CMIP6 planetary albedo climatology using satellite observations
% Satoh et al. (2019), Global Cloud-Resolving Models
% Hyder et al. (2018), Critical Southern Ocean climate model biases traced to atmospheric model cloud errors

\documentclass[12pt,a4paper]{article}
\usepackage{fontspec}
\usepackage{microtype}
\usepackage[margin=2cm]{geometry}
\usepackage[citecolor=black]{hyperref}
\usepackage[authoryear,round]{natbib}
\usepackage{doi}
\usepackage{titlesec}
\usepackage{caption}
\DeclareCaptionLabelSeparator{pipe}{ | }
\captionsetup{labelfont={bf,sf},labelsep=pipe}
\usepackage{graphicx}
\usepackage{lineno}
\linenumbers
\setsansfont{Source Sans Pro}
\setmainfont{EB Garamond}
\setmonofont[Scale=0.8]{SpaceMono}[
	Path=./fonts/,
	Extension=.ttf,
	UprightFont=*-Regular,
	BoldFont=*-Bold,
	ItalicFont=*-Italic,
	BoldItalicFont=*-BoldItalic
]
\hypersetup{
	colorlinks=true,
	urlcolor=blue,
	pdftitle={Ship-based lidar evaluation of Southern Ocean clouds in the storm-resolving general circulation model ICON, and the ERA5 and MERRA-2 reanalyses},
	pdfauthor={Peter Kuma, Frida A.-M. Bender, Adrian J. McDonald, Simon P. Alexander, Greg M. McFarquhar, John J. Cassano, Graeme E. Plank, Sean Hartery, Simon Parsons, Sally Garrett, and Alex J. Schuddeboom}
}
\bibliographystyle{abbrvnat}
\titleformat*{\section}{\large\bfseries\sffamily}
\titleformat*{\subsection}{\normalsize\bfseries\sffamily}
\titlespacing*{\section}{0pt}{14pt}{6pt}
\titlespacing*{\subsection}{0pt}{13pt}{4pt}
\parskip=6pt
\parindent=0pt

\begin{document}

\fontsize{13pt}{15pt}\selectfont

\begin{center} \Large \sffamily\textbf{Ship-based lidar evaluation of Southern Ocean clouds in the storm-resolving general circulation model ICON and the ERA5 and MERRA-2 reanalyses}\\[0.4cm]
\normalsize Peter Kuma$^\mathsf{1,2,\star}$, Frida A.-M. Bender$^\mathsf{1,2}$, Adrian J. McDonald$^\mathsf{3}$, Simon P. Alexander$^\mathsf{4,5}$, Greg M. McFarquhar$^\mathsf{6,7}$, John J. Cassano$^\mathsf{8,9,10}$, Graeme E. Plank$^\mathsf{3}$, Sean Hartery$^\mathsf{11}$, Simon Parsons$^\mathsf{12}$, Sally Garrett$^\mathsf{13}$, and Alex J. Schuddeboom$^\mathsf{3}$\\[0.4cm]
\footnotesize
$^\mathsf{1}$Department of Meteorology (MISU), Stockholm University, Stockholm, Sweden\\
$^\mathsf{2}$Bolin Centre for Climate Research, Stockholm University, Stockholm, Sweden\\
$^\mathsf{3}$School of Physical and Chemical Sciences, University of Canterbury, Christchurch, Aotearoa/New Zealand\\
$^\mathsf{4}$Australian Antarctic Division, Kingston, Tasmania, Australia\\
$^\mathsf{5}$Australian Antarctic Program Partnership, Institute for Marine and Antarctic Studies, University of Tasmania, Hobart, Tasmania, Australia\\
$^\mathsf{6}$Cooperative Institute of Severe and High Impact Weather Research and Operations, University of Oklahoma, Norman, OK, USA\\
$^\mathsf{7}$School of Meteorology, University of Oklahoma, Norman, OK, USA\\
$^\mathsf{8}$Cooperative Institute for Research in Environmental Sciences, University of Colorado, Boulder, CO, USA\\
$^\mathsf{9}$National Snow and Ice Data Center, University of Colorado, Boulder, CO, USA\\
$^\mathsf{10}$Department of Atmospheric and Oceanic Sciences, University of Colorado, Boulder, CO, USA\\[0.2cm]
$^\mathsf{11}$Department of Physics \& Atmospheric Science, Dalhousie University, Halifax, Canada\\
$^\mathsf{12}$New South Wales Department of Planning and Environment, Sydney, New South Wales, Australia\\
$^\mathsf{13}$New Zealand Defence Force, Wellington, New Zealand\\
$^\star$Corresponding author: Peter Kuma (\href{mailto:peter.kuma@misu.su.se}{peter.kuma@misu.su.se})\\[0.4cm]
\normalsize \today\\[0.4cm]

\end{center}

\section*{Abstract}

Global storm-resolving models (GSRMs) are the next avenue of climate modelling.
Among them is a 5-km version of the Icosahedral Nonhydrostatic Weather and
Climate Model (ICON). The high resolution means that parameterizations of
convection and clouds including subgrid-scale clouds are forgone, relying on
explicit their simulation, but still using microphysics and turbulence
parametrization.  Standard-resolution (10--100-km) models, which use
subgrid-scale convection and cloud parametrization, have substantial cloud
biases over the Southern Ocean (SO), affecting radiation and sea surface
temperature.  We evaluated SO clouds in ICON and the ERA5 and MERRA-2
reanalyses.  The SO is dominated by low clouds, which cannot be observed
accurately from space due to overlapping clouds, attenuation, and ground
clutter.  Instead, we analysed about 2400 days of lidar observations from 31
voyages and a station during 2010--2021 using a ground-based lidar simulator.
ICON and the reanalyses underestimate the total cloud fraction by about 10 and
20\%, respectively. ICON and ERA5 overestimate the cloud occurrence peak at
about 500 m, potentially explained by their lifting condensation levels being
too high.  The reanalyses strongly underestimate near-surface clouds or fog.
MERRA-2 tends to underestimate cloud occurrence at all heights.  In daily cloud
cover, ICON and the reanalyses tend to be about 1 and 2 oktas clearer,
respectively.  Conditions of weak stability are associated with the largest
errors in ICON and the reanalyses.  Compared to radiosondes, potential
temperature is accurate in the reanalyses to about 0.5 K, but ICON
underestimates stability over the low-latitude SO and too humid in the boundary
layer. MERRA-2 is too humid at all heights. SO cloud biases remain a
substantial issue in the GSRM, but are an improvement over the lower-resolution
reanalyses. Explicitly resolved convection and cloud processes (except
microphysics) were not enough to address the model cloud biases.

\section{Introduction}
\label{sec:introduction}

Increasing climate model resolution is one way of improving model accuracy of
representation of the climate system \citep{mauritsen2022}. It has been
practiced since the advent of climate modelling as more computational power,
memory, and storage capacity become available. It is, however, often not as
easy as changing the grid size because of the complex interplay between model
dynamics and physics, which necessitates adjusting and tuning all components
together.  Increasing resolution is of course limited by the available
computational power and a trade-off with increasing parameterization
complexity, which is another way of improving model accuracy.  Current
computational availability and acceleration from general-purpose computing on
graphics processing units (GPUs) is progressing to enable km-scale (also called
k-scale) Earth system models (ESMs) and coupled atmosphere--ocean general
circulation models (AOGCMs) in research conditions today and operationally in
the forthcoming years.  Therefore, it represents a natural advance in climate
modelling.  Global storm-resolving models (GSRMs) are emerging as a new front
in the development of high-resolution global climate models, with horizontal
grid resolutions of about 2--8 km \citep{satoh2019,stevens2019}. This is enough
to resolve mesoscale convective storms, but smaller-scale convective plumes and
cloud structure remain unresolved. At an approximately 5-km scale,
non-hydrostatic processes also become important \citep{weisman1997}, and for
this reason such models are generally non-hydrostatic. The terms global
cloud-resolving models or global convection-permitting/-resolving models are
also sometimes used interchangeably with GSRMs but imply that clouds or
convection are resolved explicitly, which is not entirely true for GSRMs, as
this would require an even higher horizontal resolution \citep{satoh2019}.
Representative of these efforts is the DYnamics of the Atmospheric general
circulation Modeled On Non-hydrostatic Domains (DYAMOND) project
\citep{stevens2019,dyamond}, which is an intercomparison of nine global GSRMs
over two 40-day time periods in summer (1 August -- 10 September 2016) and
winter (20 January -- 1 March 2020). A new one-year GSRM intercomparison is
currently proposed by \cite{takasuka2024}, with the hope of also evaluating the
seasonal cycle and large-scale circulation.  An alternative to using a
computationally costly GSRM is to train an artificial neural network on GSRM
output and use it for subgrid-scale clouds, as done with the GSRM ICON by
\cite{grundner2022} and \cite{grundner2023}.

nextGEMS is a project \citep{nextgems} focused on the research and development
of GSRMs at multiple modelling centres and universities in Europe.  The project
also develops GSRM versions of the Icosahedral Nonhydrostatic Weather and
Climate Model (ICON; \cite{hohenegger2023}), the Integrated Forecasting System
[IFS; \cite{ifs2023}], and their ocean components at eddy-resolving
resolutions: ICON-O \citep{korn2022} coupled with ICON and
Finite-Element/volumE Sea ice-Ocean Model [FESOM; \cite{wang2014}] and Nucleus
for European Modelling of the Ocean [NEMO; \cite{madec2023}] coupled with IFS.
The project has so far produced ICON and IFS simulations with three development
versions called Cycle 1--3 and a pre-final version, with a final production
version planned by the end of the project. nextGEMS is not the only project
developing GSRMs; other GSRMs (or GSRM versions of climate models) currently in
development include: Convection-Permitting Simulations With the E3SM Global
Atmosphere Model [SCREAM; \cite{caldwell2021}], Atmospheric Model [NICAM;
\cite{satoh2008}], Unified Model (UM), eXperimental System for High-resolution
modeling for Earth-to-Local Domain [X-SHiELD; \cite{shield}], Action de
Recherche Petite Echelle Grande Echelle-NonHydrostatic version [ARPEGE-NH;
\cite{bubnova1995,voldoire2017}], Finite-Volume Dynamical Core on the Cubed
Sphere [FV3, \cite{lin2004}], the National Aeronautics and Space Administration
(NASA) Goddard Earth Observing System global atmospheric model version 5
[GEOS5; \cite{putman2011}], Model for Prediction Across Scales [MPAS;
\cite{skamarock2012}], and System for Atmospheric Modeling [SAM;
\cite{khairoutdinov2003}].

Multiple cloud properties have an effect on shortwave (SW) and longwave (LW)
radiation. To first order, the total cloud fraction, cloud phase, and the
liquid and ice water path are the most important cloud properties influencing
SW and LW radiation. These properties are  in turn influenced by the
atmospheric thermodynamics, convection and circulation, and both the indirect
and direct effects of aerosols. Second order effects on SW and LW radiation are
associated with the cloud droplet size distribution, ice crystal habit, cloud
lifetime, and direct radiative interaction with aerosols.  In the
6$^\mathrm{th}$ phase of the Coupled Model Intercomparison Project [CMIP6;
\cite{eyring2016}], the cloud feedback has increased relative to CMIP5
\citep{zelinka2020}, which is one of the main reasons for the higher climate
sensitivity of CMIP6 models.

The Southern Ocean (SO) is known to be a problematic region for climate model
biases due to a lack of surface and in situ observations and being a lower
priority region for numerical weather prediction (NWP) and climate model
development because of its distance from populated areas.  Nevertheless,
radiation biases and changes over an area of its size have a substantial
influence on the global climate, such as affecting the Earth radiation balance
and carbon cycle, and the SO is an important part of the global ocean conveyor
belt. In general, marine clouds have a disproportionate effect on top of
atmosphere (TOA) SW radiation due to the relatively low albedo of the sea
surface.  The relative longitudinal symmetry of the SO means that model cloud
biases tend to be similar across longitudes.  Here, we conventionally refer to
the SO as ocean regions south of 40°S, low-latitude SO as 40--55°S and
high-latitude SO as south of 55°S. The reason for this dividing latitude
is to split the SO into about two equal zones, as well as the results by
\cite{schuddeboom2021} (Fig. 2b), showing a contrast in CMIP model radiation
biases, \citep{schuddeboom2019} (Fig. 2) and \cite{kuma2020} (Fig. 3) showing
contrasting radiation biases in the Hadley Centre Global Environmental Model
between these two zones. The findings of \cite{niu2024}, however, support a
different dividing line of 62º based on cloud condensation nuclei
concentration.

SO radiation biases have been relatively large and systematic compared to the
rest of the globe since at least CMIP3 \citep{trenberth2010}, and the SO SW
cloud radiative effect (CRE) bias is still positive in eight analysed CMIP6
models analysed by \cite{schuddeboom2021} over the high-latitude SO, whereas
over the low-latitude SO it tends to be more neutral or negative in some
models. Too much absorbed SW radiation over the SO was also identified in the
GSRM SCREAM \cite{caldwell2021}. Compensating biases are possible, such as the
`too few too bright' cloud bias, characterised by too small cloud fraction and
too large cloud albedo \citep{wall2017,kuma2020}, previously described by
\cite{webb2001}, \cite{weare2004}, \cite{zhang2005}, \cite{karlsson2008},
\cite{nam2012}, \cite{klein2013}, and \cite{bender2017} in other regions and
models, which means that a model can maintain a reasonable SW radiation balance
by reflecting too much SW radiation from clouds, but these cover too small
area.  A study by \cite{konsta2022} showed that this type of bias is still
present in six analysed CMIP6 models in tropical marine clouds, using the
GCM-Oriented Cloud-Aerosol Lidar and Infrared Pathfinder Satellite Observation
(CALIPSO) Cloud Product [CALIPSO--GOCCP; \cite{chepfer2010}] and Polarization
\& Anisotropy of Reflectances for Atmospheric Sciences coupled with
Observations from a Lidar [PARASOL; \cite{lier2008}] as a reference. They
suggest improper simulation of subgrid-scale cloud heterogeneity as a cause.
Compensating cloud biases in the Australian Community Climate and Earth System
Simulator (ACCESS) – Atmosphere-only model version 2 (AM2) over the SO were
analysed by \cite{fiddes2022} and \cite{fiddes2024}.  \cite{possner2022} showed
that over the SO, the DYAMOND GSRM ICON underestimates low-level cloud fraction
on the order of 30\% [relative to Moderate Resolution Imaging Spectroradiometer
(MODIS) data] and overestimates downwelling TOA SW radiation by approximately
10 Wm$^\mathrm{-2}$ [relative to the Clouds and the Earth’s Radiant Energy
System (CERES) data] in the highest model resolution run (2.5 km).
\cite{zhao2022} reported a similar SW radiation bias in five analysed CMIP6
models over high-latitude SO and the total cloud fraction underestimation on
the order of 10\% over the entire 40--60°S SO. Recently, \cite{ramadoss2024}
analysed 48 hours of km-scale ICON limited area model NWP simulations over a SO
region adjacent to Tasmania against the Clouds, Aerosols, Precipitation,
Radiation, and atmospherIc Composition Over the southeRn oceaN (CAPRICORN)
voyage cloud and precipitation observations \cite{mcfarquhar2021}. They found
the ICON cloud optical thickness was underestimated relative to Himawari‐8
satellite observations, but also identified large differences in cloud top
phase.

In general, sea surface temperature (SST) biases in the SO can originate either
in the atmosphere \citep{hyder2018}, caused by too much shortwave heating of
the surface, too little longwave cooling of the surface, or in the ocean
circulation.  Interactions of both are also possible, for example SST affecting
clouds and clouds affecting the surface radiation. Using the European Centre
for Medium-Range Weather Forecasts (ECMWF) Reanalysis 5 (ERA5) as a reference,
\cite{zhang2023} have shown that SST biases have improved in CMIP6 compared to
CMIP5, with SST overall increasing in the later CMIP phase. However, over the
SO this resulted in an even higher positive bias, especially in the Atlantic
Ocean (AO) sector of the SO, increasing by up to 1°C.  \cite{luo2023}
identified that the SO SST bias in an ensemble of 18 CMIP6 models originates
not from the surface heat and radiation fluxes (using reanalyses as a
reference), but from a warm bias in the Northern Atlantic Deep Water.

The main aim of this study is to evaluate the GSRM version of ICON. ICON
developed and maintained jointly by Deutscher Wetterdienst,
Max-Planck-Institute for Meteorology, Deutsches Klimarechenzentrum (DKRZ),
Karlsruhe Institute of Technology, and the Center for Climate Systems Modeling.
Previous studies have identified substantial large-scale biases in climate
model clouds over the SO, affecting sea surface temperature and the Earth’s
albedo. Our aim is to quantify how well the GSRM ICON is simulating clouds in
this region, particularly in light of the fact that subgrid-scale clouds and
convection are not parameterized in this model. This region is mostly dominated
by boundary layer clouds generated by shallow convection, and these are
problematic to observe by spaceborne lidars and radars, which are affected by
attenuation by overlapping and thick clouds and ground clutter, respectively.
Specifically, the radar on CloudSat and lidar on CALIPSO (neither of which are
nowoperational) are affected by the abovementioned issues, resulting in a
strong underestimation of cloud occurrence below 2 km relative to ground-based
lidar observations \citep{mcerlich2021}.  We hypothesise that this, in turn,
can lead to systematic biases in low clouds in climate models, which are
frequently evaluated against CloudSat--CALIPSO products. Reanalyses can also
suffer from cloud biases, as these are usually parametrised in their
atmospheric component, and also in regions where input observations are sparse.
This makes them a problematic reference for clouds over the SO, and any biases
relative to a reanalysis should be interpreted with caution. Instead, we chose
to use a large set of ship-based observations conducted with ceilometers and
lidars on board of the RV \emph{Polarstern} and other voyages and stations as a
reference for the model evaluation.

Altogether, we analysed about 2400 days of data from 31 voyages, and one
sub-Antarctic station covering diverse longitudes and latitudes of the SO. To
achieve a like-for-like comparison with the model, we used a ground-based lidar
simulator called the Automatic Lidar and Ceilometer Framework [ALCF;
\cite{kuma2021}]. We contrasted the results with ERA5 \citep{era5} and the
Modern-Era Retrospective analysis for Research and Applications, Version 2
[MERRA-2; \cite{gelaro2017}].

\begin{figure}[b!]
\centering
\includegraphics[width=\textwidth]{img/map_fig.pdf}
\caption{
\textbf{(a)} A map showing the tracks of 31 voyages of RV \emph{Polarstern},
RSV \emph{Aurora Australis}, RV \emph{Tangaroa}, RV \emph{Nathaniel B. Palmer},
and HMNZS \emph{Wellington} and one sub-Antarctic station (Macquarie Island)
analysed here. The tracks cover Antarctic sectors south of South America, the
Atlantic Ocean, Africa, Australia, and New Zealand in the years 2010--2021
(inclusive).  The dotted and dashed lines at 40°S and 55°S delineate the
Southern Ocean area of our analysis and its partitioning into two subsets,
respectively.  A photo of \textbf{(b)} RV \emph{Polarstern} (© Folke Mehrtens,
Alfred-Wegener-Institut), \textbf{(c)} Lufft CHM 15k installed on RV
\emph{Tangaroa} (© Peter Kuma, University of Canterbury), \textbf{(d)} Vaisala
CL51 (© Jeff Aquilina, Bureau of Meteorology), \textbf{(e)} Vaisala CT25K at
Macquarie Island (© Simon P.\ Alexander, Australian Antarctic Division).
}
\label{fig:map}
\end{figure}

\section{Methods}
\label{sec:methods}

\subsection{Voyage and station data}

Together, we analysed data from 31 voyages of RV \emph{Polarstern}, the
resupply vessel (RSV) \emph{Aurora Australis}, RV \emph{Tangaroa}, RV
\emph{Nathaniel B.  Palmer}, Her (now His) Majesty's New Zealand Ship (HMNZS)
\emph{Wellington} and one sub-Antarctic station (Macquarie Island) in the SO
south of 40°S between 2010 and 2021. Fig. \ref{fig:map} shows a map of the
voyages, Table \ref{tab:voyages} list the voyages, campaigns, and stations, and
Table \ref{tab:voyage-references} lists references where available. Altogether,
the voyages and station dataset comprised 2421 days of data south of 40°S, but
the availability of ceilometer data was slightly shorter due to gaps in
measurements.

Missing days in the ceilometer data were HMNZSW16 (7 days): 24--27 November, 10
December, 16--17 December 2016; Measurements of Aerosols, Radiation, and CloUds
over the Southern Ocean (MARCUS;  3 days): 8, 10 November, 10 December 2017;
Macquarie Island Cloud Radiation Experiment (MICRE; 9 days): 7--8, 29 June, 5,
16 July, 15 August, 17 October 2016, 11 February, 21 March 2017; TAN1502 (1
day): 24 January.

The data sources contained ceilometer observations captured by the Vaisala CL51
and CT25K, and the Lufft CHM 15k, described in detail below (Sections
\ref{sec:cl51} and \ref{sec:chm15k}). A ceilometer is a low-power near-infrared
vertically pointing lidar principally designed to  measure cloud base, but they
also measure the full vertical structure of clouds as long as the laser signal
is not attenuated by thick clouds, which can be used to infer additional
information such as a cloud mask and cloud occurrence by height.  We note that
during the MICRE campaign, the ceilometers Vaisala CT25K and CL51 were
installed at the Macquarie Island station concurrently, but in our analysis we
only used the CT25K data obtained from the Atmospheric Radiation Measurement
(ARM) data archive.

Apart from lidar observations, radiosondes were launched on weather balloons at
regular synoptic times on the RV \emph{Polarstern}, MARCUS, NBP17024, TAN1702,
and TAN1802 voyages and campaigns, measuring pressure, temperature, relative
humidity, and the global navigation satellite system coordinates. Derived
thermodynamic (virtual potential temperature, lifting condensation level, etc.)
and dynamic physical quantities (wind speed and direction) for the measured
vertical profiles were calculated with rstool \citep{rstool}. Surface
meteorological quantities were measured continuously by an onboard automatic
weather station or individual instruments.

\begin{table}[p!]
\caption{
An overview of the analysed voyages, campaigns, and stations. Start, end, and
the number of days (UTC; inclusive) refer to the time period when the vessel
was south of 40°S.  Abbreviations: ceilometer (ceil.), Australia (AU), New
Zealand (NZ), South America (SA), Atlantic Ocean (AO), and Africa (AF). The
number of days is rounded to the nearest integer. CL51/31 indicates CL51
configured to emulate CL31.
}
\label{tab:voyages}
\small
\begin{tabular}{llllllr}
\textbf{Name} & \textbf{Vessel or station} & \textbf{Ceil.} & \textbf{Region} & \textbf{Start} & \textbf{End} & \textbf{Days}\\
\hline
AA15-16  & RSV \emph{Aurora Australis}   & CL51    & AU       & 2015-10-22 & 2016-02-22 & 124 \\
HMNZSW16 & HMNZS \emph{Wellington}       & CHM 15k & NZ       & 2016-11-23 & 2016-12-19 & 27 \\
MARCUS   & RSV \emph{Aurora Australis}   & CT25K   & AU       & 2017-10-29 & 2018-03-26 & 149 \\
MICRE    & Macquarie Is. station         & CT25K   & AU/NZ    & 2016-04-03 & 2018-03-14 & 710 \\
NBP1704  & RV \emph{Nathaniel B. Palmer} & CHM 15k & NZ       & 2017-04-14 & 2017-06-08 & 55 \\
PS77/2   & RV \emph{Polarstern}          & CL51    & SA/AO/AF & 2010-12-01 & 2011-02-04 & 65 \\
PS77/3   & RV \emph{Polarstern}          & CL51    & SA/AO/AF & 2011-02-07 & 2011-04-14 & 66 \\
PS79/2   & RV \emph{Polarstern}          & CL51    & SA/AO/AF & 2011-12-06 & 2012-01-02 & 27 \\
PS79/3   & RV \emph{Polarstern}          & CL51    & SA/AO/AF & 2012-01-10 & 2012-03-10 & 61 \\
PS79/4   & RV \emph{Polarstern}          & CL51    & SA/AO/AF & 2012-03-14 & 2012-04-08 & 26 \\
PS81/2   & RV \emph{Polarstern}          & CL51    & SA/AO/AF & 2012-12-02 & 2013-01-18 & 47 \\
PS81/3   & RV \emph{Polarstern}          & CL51    & SA/AO/AF & 2013-01-22 & 2013-03-17 & 55 \\
PS81/4   & RV \emph{Polarstern}          & CL51    & SA/AO/AF & 2013-03-18 & 2013-04-16 & 30 \\
PS81/5   & RV \emph{Polarstern}          & CL51    & SA/AO/AF & 2013-04-20 & 2013-05-23 & 33 \\
PS81/6   & RV \emph{Polarstern}          & CL51    & SA/AO/AF & 2013-06-10 & 2013-08-12 & 63 \\
PS81/7   & RV \emph{Polarstern}          & CL51    & SA/AO/AF & 2013-08-15 & 2013-10-14 & 60 \\
PS81/8   & RV \emph{Polarstern}          & CL51    & SA/AO/AF & 2013-11-12 & 2013-12-14 & 31 \\
PS81/9   & RV \emph{Polarstern}          & CL51    & SA/AO/AF & 2013-12-21 & 2014-03-02 & 71 \\
PS89     & RV \emph{Polarstern}          & CL51    & SA/AO/AF & 2014-12-05 & 2015-01-30 & 56 \\
PS96     & RV \emph{Polarstern}          & CL51    & SA/AO/AF & 2015-12-08 & 2016-02-14 & 68 \\
PS97     & RV \emph{Polarstern}          & CL51    & SA/AO/AF & 2016-02-15 & 2016-04-06 & 52 \\
PS103    & RV \emph{Polarstern}          & CL51    & SA/AO/AF & 2016-12-18 & 2017-02-02 & 46 \\
PS104    & RV \emph{Polarstern}          & CL51    & SA/AO/AF & 2017-02-08 & 2017-03-18 & 39 \\
PS111    & RV \emph{Polarstern}          & CL51    & SA/AO/AF & 2018-01-21 & 2018-03-14 & 52 \\
PS112    & RV \emph{Polarstern}          & CL51    & SA/AO/AF & 2018-03-18 & 2018-05-05 & 49 \\
PS117    & RV \emph{Polarstern}          & CL51    & SA/AO/AF & 2018-12-18 & 2019-02-07 & 51 \\
PS118    & RV \emph{Polarstern}          & CL51    & SA/AO/AF & 2019-02-18 & 2019-04-08 & 50 \\
PS123    & RV \emph{Polarstern}          & CL51    & SA/AO/AF & 2021-01-10 & 2021-01-31 & 21 \\
PS124    & RV \emph{Polarstern}          & CL51    & SA/AO/AF & 2021-02-03 & 2021-03-30 & 55 \\
TAN1502  & RV \emph{Tangaroa}            & CL51/31 & NZ       & 2015-01-20 & 2015-03-12 & 51 \\
TAN1702  & RV \emph{Tangaroa}            & CHM 15k & NZ       & 2017-03-09 & 2017-03-31 & 23 \\
TAN1802  & RV \emph{Tangaroa}            & CHM 15k & NZ       & 2018-02-07 & 2018-03-20 & 41 \\
\hline
\textbf{Total} &                         &         &          &            &            & \textbf{2421}\\
\end{tabular}
\normalsize
\end{table}

\begin{table}[t!]
\caption{Voyage, campaign and station publication references.}
\label{tab:voyage-references}
\small
\begin{tabular}{lp{14.5cm}}
\textbf{Name} & \textbf{References}\\
\hline
AA15-16  & \cite{klekociuk2020} \\
MARCUS   & \cite{mcfarquhar2021,xia2024,niu2024} \\
MICRE    & \cite{mcfarquhar2021} \\
NBP1704  & \cite{ackley2020} \\
PS77/2   & \cite{kniglanglo2011a,kniglanglo2011b,kniglanglo2011c,kniglanglo2014a,fahrbach2011} \\
PS77/3   & \cite{kniglanglo2011d,kniglanglo2011e,kniglanglo2012a,kniglanglo2014b,knust2011} \\
PS79/2   & \cite{kniglanglo2012b,kniglanglo2012c,kniglanglo2012d,kniglanglo2014c,kattner2012} \\
PS79/3   & \cite{kniglanglo2012e,kniglanglo2012f,kniglanglo2012g,kniglanglo2014d,wolfgladrow2012} \\
PS79/4   & \cite{kniglanglo2012h,kniglanglo2012i,kniglanglo2012j,kniglanglo2014e,lucassen2012} \\
PS81/2   & \cite{kniglanglo2013a,kniglanglo2013b,kniglanglo2013c,kniglanglo2014f,boebel2013} \\
PS81/3   & \cite{kniglanglo2013d,kniglanglo2013e,kniglanglo2013f,kniglanglo2014g,gutt2013} \\
PS81/4   & \cite{kniglanglo2013g,kniglanglo2013h,kniglanglo2013i,kniglanglo2014f,bohrmann2013} \\
PS81/5   & \cite{kniglanglo2013j,kniglanglo2013k,kniglanglo2013l,kniglanglo2014g,jokat2013} \\
PS81/6   & \cite{kniglanglo2013m,kniglanglo2013n,kniglanglo2013o,kniglanglo2014h,lemke2013} \\
PS81/7   & \cite{kniglanglo2013p,kniglanglo2013q,kniglanglo2014i,kniglanglo2016a,meyer2013} \\
PS81/8   & \cite{kniglanglo2013r,kniglanglo2014j,kniglanglo2014k,kniglanglo2014l,schlindwein2014} \\
PS81/9   & \cite{kniglanglo2014m,kniglanglo2014n,kniglanglo2014o,kniglanglo2014p,knust2014} \\
PS89     & \cite{kniglanglo2015a,kniglanglo2015b,kniglanglo2015c,kniglanglo2015d,boebel2016}\\
PS96     & \cite{kniglanglo2016b,kniglanglo2016c,kniglanglo2016d,kniglanglo2016e,schrder2017} \\
PS97     & \cite{kniglanglo2016f,kniglanglo2016g,kniglanglo2016h,kniglanglo2016i,lamy2017} \\
PS103    & \cite{kniglanglo2017a,kniglanglo2017b,kniglanglo2017c,kniglanglo2017d,boebel2018} \\
PS104    & \cite{kniglanglo2017e,kniglanglo2017f,kniglanglo2017g,gohl2018,schmithsen2021a} \\
PS111    & \cite{schmithsen2019a,schmithsen2020a,schmithsen2021b,schmithsen2021c,schrder2018} \\
PS112    & \cite{schmithsen2019b,schmithsen2020b,schmithsen2021d,schmithsen2021e,meyer2018} \\
PS117    & \cite{schmithsen2019c,schmithsen2020c,schmithsen2021f,schmithsen2021g,boebel2019} \\
PS118    & \cite{schmithsen2019d,schmithsen2020d,schmithsen2021h,schmithsen2021i,dorschel2019} \\
PS123    & \cite{schmithsen2021j,schmithsen2021m,schmithsen2021n,schmithsen2021k,hoppmann2023} \\
PS124    & \cite{schmithsen2021o,schmithsen2021q,schmithsen2021p,hoppmann2023} \\
TAN1802  & \cite{kremser2020,kremser2021} \\
\end{tabular}
\end{table}

\subsection{Vaisala CL51 and CT25K}
\label{sec:cl51}

The Vaisala CL51 and CT25K (photos in Fig. \ref{fig:map}d, e) are ceilometers
operating at a near-infrared wavelength of 910 nm and 905 nm, respectively.
The CL51 can also be configured to emulate the Vaisala CL31. The maximum range
is 15.4 km (CL51), 7.7 km (CL31 emulation mode with 5 m vertical resolution),
and 7.5 km (CT25K). The vertical resolution is 10 m (5 m configurable) in CL51
and 30 m in CT25K observations. The sampling (temporal) resolution is
configurable, and in our datasets is approximately 6 s for CL51 on AA15‐16, 16
s for CT25K on MARCUS and MICRE, 36 s for CL51 on RV \emph{Polarstern}, and
about 2.37 s for CL51 with CL31 emulation on TAN1502. The wavelengths of 905
and 910 nm are affected by water vapour absorption of about 20\% in the
mid-latitudes \citep{wiegner2015,wiegner2019}, with 910 nm affected more
strongly, but we do not expect this to be a significant issue as explained in
\cite{kuma2021}.  The instrument data files containing raw uncalibrated
backscatter were first converted to Network Common Data Form (NetCDF) with
cl2nc (\url{https://github.com/peterkuma/cl2nc}) and then processed with the
ALCF (Section \ref{sec:alcf}) to produce absolutely calibrated attenuated
volume backscattering coefficient (AVBC), cloud mask, cloud occurrence by
height, and the total cloud fraction. Because the CT25K uses a very similar
wavelength to CL51, equivalent calculations as for CL51 were done assuming a
wavelength of 910 nm. The Vaisala CL51 and CT25K instruments were used on most
of the voyages and stations analysed here. Fig.  \ref{fig:example}a shows an
example of AVBC derived from the CL51 instrument data.

\begin{figure}[b!]
\centering
\includegraphics[width=\textwidth]{img/example.png}
\caption{
An example of attenuated volume backscattering coefficient (AVBC) \textbf{(a)}
measured by CL51 during 24 hours on the PS81/3 voyage and \textbf{(b)} an
equivalent AVBC simulated with the ALCF from ERA5 data during the same time
period. The red line identifies the cloud mask determined by the ALCF.
}
\label{fig:example}
\end{figure}

\subsection{Lufft CHM 15k}
\label{sec:chm15k}

The Lufft CHM 15k (photo in Fig. \ref{fig:map}c) is a ceilometer operating at a
near-infrared wavelength of 1064 nm. The maximum range is 15.4 km, the vertical
resolution is 5 m in the near range (up to 150 m) and 15 m above, the sampling
(temporal) resolution is 2 s, and the number of vertical levels is 1024.
NetCDF files containing uncalibrated backscatter produced by the instrument
were processed with the ALCF (Section \ref{sec:alcf}) to again produce AVBC,
cloud mask, cloud occurrence by height, and the total cloud fraction. The CHM
15k was used on four voyages (HMNZSW16, TAN1702, TAN1802, and NBP1704).

\subsection{ALCF}
\label{sec:alcf}

The Automatic Lidar and Ceilometer Framework (ALCF) is a ground-based lidar
simulator and a tool for processing observed lidar data, supporting various
instruments and models \citep{kuma2021}. It performs radiative transfer
calculations to derive equivalent lidar AVBC in an atmospheric model, which can
then be compared with observed AVBC. For this purpose, it takes the cloud
fraction, liquid and ice mass mixing ratio, temperature, and pressure model
fields as an input and is run offline (on the model output rather than inside
the model code). The lidar simulator in the ALCF is based on the instrument
simulator Cloud Feedback Model Intercomparison Project (CFMIP) Observation
Simulator Package (COSP) \citep{bodas-salcedo2011}.  After AVBC is calculated,
a cloud mask, cloud occurrence by height, and the total cloud fraction are
determined. The ALCF has been used by several research teams for model and
reanalysis evaluation
\citep{kuma2020,kremser2021,guyot2022,pei2023,whitehead2023,mcdonald2024}.

Absolute calibration of the observed backscatter was performed by comparing the
measured clear-sky molecular backscatter statistically with simulated clear-sky
molecular backscatter. AVBC was resampled to 5 min temporal resolution and 50 m
vertical resolution to increase signal-to-noise ratio while having enough
resolution to detect small-scale cloud variability. The noise standard
deviation was calculated from AVBC at the highest range, where no clouds are
expected.  A cloud mask was calculated from AVBC using a fixed threshold of
$\mathrm{2\times 10^{-6} m^{-1}sr^{-1}}$ after subtracting 5 standard
deviations of range-scaled noise. Fig. \ref{fig:example}b shows an example of
simulated Vaisala CL51 backscatter from ERA5 data, corresponding to a day of
measurements by the instrument on the PS81/3 voyage.

\subsection{ICON}
\label{sec:icon}

A coupled (atmosphere--ocean) GSRM version of the ICON model is in development
at the nextGEMS project \citep{hohenegger2023}. ICON is an exceptionally
versatile model, allowing for simulations ranging from coarse-resolution ESM
simulations, GSRM simulations, limited area model simulations, to large eddy
simulations (LES), for both weather prediction and climate projections. ICON
uses the atmospheric component ICON-A \citep{giorgetta2018}, whose physics is
derived from ECHAM6 \citep{stevens2013}, and the ocean component ICON-O
\citep{korn2022}. Earlier runs of the GSRM ICON from DYAMOND were evaluated by
\cite{mauritsen2022}.

Here, we use a free-running (i.e., the weather situation in the model does not
correspond to reality) coupled GSRM simulation made for the purpose of climate
projection.  nextGEMS has so far produced four cycles of model runs. We used a
Cycle 3 run \emph{ngc3028} produced in 2023 \citep{nextgems2023a,nextgems2023b}
for a model time period of 20 January 2020 to 22 July 2025, of which we
analysed the period 2021--2024 (inclusive). While a Cycle 4 run was available,
we could not use it due to a lack of availability of the necessary variables.
The horizontal resolution of ngc3028 is about 5 km.  The model output is
available on 90 vertical levels and 3-hourly instantaneous temporal resolution.

Unlike current general circulation models (GCMs), the storm-resolving version
of ICON does not use convective and cloud parameterization but relies on
explicit simulation of convection and clouds on the model grid. Subgrid-scale
clouds are not resolved, and the grid cell cloud fraction is always either 0 or
100\%. While this makes the code development simpler without having to rely on
uncertain parameterizations, it can miss smaller-scale clouds below the grid
resolution.  Turbulence and cloud microphysics are still parameterized in this
model. To account for the radiative effects of subgrid-scale clouds, a cloud
inhomogeneity factor is introduced in the model which scales down the cloud
liquid water for radiative calculations. It ranges from 0.4 at lower
tropospheric stability (LTS) of 0 K to 0.8 at 30 K. In addition, vertical
mixing is enhanced in unstable and humid lower-tropospheric conditions, which
reduces the amount of shallow clouds.

Because the model is free-running, weather and climate oscillations (such as
the El Niño--Southern Oscillation) are not expected to be equivalent to reality
at the same time and place. To compare with the observations collected in
different years (2010--2021, inclusive), we compared the model output with
observations at the same time of year and geographical location, as determined
for each data point such as a lidar profile or a radiosonde launch.

\subsection{MERRA-2}

The Modern-Era Retrospective analysis for Research and Applications, Version 2
(MERRA-2) is a reanalysis produced by the Global Modeling and Assimilation
Office at the NASA Goddard Space Flight Center \citep{gelaro2017}.  It uses
version 5.12.4 of the Goddard Earth Observing System (GEOS) atmospheric model
\citep{rienecker2008,molod2015}. Non-convective clouds (condensation,
autoconversion, and evaporation) are parametrised using a prognostic scheme
\citep{bacmeister2006}, and sub-grid cloud fraction is determined using total
water distribution and a critical relative humidity threshold. The reanalysis
output analysed here is available at a spatial resolution of 0.5° of latitude
and 0.625° of longitude, which is about 56 km in the North--South direction and
35 km in the East--West direction at 60°S. The number of vertical model levels
is 72. Here, we use the following products: 1-hourly instantaneous 2D
single-level diagnostics (M2I1NXASM) for 2-m temperature and humidity; 3-hourly
instantaneous 3D assimilated meteorological fields (M2I3NVASM) for cloud
quantities, pressure, and temperature; 1-hourly average 2D surface flux
diagnostics (M2T1NXFLX) for precipitation; and 1-hourly average 2D radiation
diagnostics (M2T1NXRAD) for radiation quantities \citep{merra2}.

\subsection{ERA5}

ERA5 \citep{era5} is a reanalysis produced by the ECMWF.  It is based on a
numerical weather prediction model IFS version CY41R2.  It uses the
\cite{tiedtke1993} prognostic cloud scheme and \cite{forbes2014} for
mixed-phase clouds. The horizontal resolution is 0.25° in latitude and
longitude, which is about 28 km in the North--South direction and 14 km in the
East--West direction at 60°S.  Internally, the model uses 137 vertical levels.
Here, we use output at 1-hourly instantaneous time intervals, except for
radiation quantities, which are accumulations (from these we calculate daily
means).  Vertically resolved quantities are made available on 37 pressure
levels.

\subsection{CERES}

TOA radiation quantities are taken from the CERES instruments onboard the Terra
and Aqua satellites \citep{wielicki1996,loeb2018}. In our analysis we used the
adjusted all sky SW and LW upwelling fluxes at TOA from the synoptic TOA and
surface fluxes and clouds 1 degree daily edition 4A product
(CER\_SYN1deg-Day\_Terra-Aqua-MODIS\_Edition4A)
\citep{doelling2013,doelling2016}.

Radiation calculations presented in the results (Section \ref{sec:results})
were completed such that they always represent daily means in order to be
consistent with the CERES SYN1deg data. Therefore, every instantaneous profile
in the simulated lidar data was assigned a daily mean radiation value
corresponding to the day (in the Coordinated Universal Time; UTC). In turn, the
average radiation during the entire voyage or station observation period were
calculated as the average of the profile values. In the observed lidar data,
the daily mean radiation value was taken from the spatially and temporally
co-located CERES SYN1deg data for the day (in UTC). The voyage or station
average was calculated in the same way.

\begin{figure}[b!]
\centering
\includegraphics[width=\textwidth]{img/ann.pdf}
\caption{
Artificial neural network (ANN) for prediction of precipitation in lidar
backscatter. \textbf{(a)} Diagram showing the TensorFlow structure of the ANN,
\textbf{(b)} randomly selected example samples of near-surface backscatter in
four categories (clear, fog, rain, and snow), as determined by coincident
human-performed weather observations, \textbf{(c)} receiver operating
characteristic diagram of the ANN, \textbf{(d)} examples of 10-day time series
of human-observed (`HUM') and predicted precipitation based on an ANN trained
on all voyages (`ANN') and all voyages except for the shown voyage (`ANN2')
during three randomly selected voyages with the available data. Here, by
`randomly selected' we mean selected from the top of a permutation generated by
a pseudo-random number generator to prevent authors' bias in the selection.
}
\label{fig:ann}
\end{figure}

\subsection{Precipitation identification using machine learning}
\label{sec:ann}

Precipitation can cause strong enough lidar backscattering to be recognised as
clouds by the threshold-based cloud detection method used in the ALCF. This is
undesirable if equivalent precipitation backscatter is not included in the
simulated lidar profiles. It was not possible to include precipitation
simulation in the ALCF due to the absence of required fields in the ICON model
output and the reanalysis data (the liquid and ice precipitation mass mixing
ratios). The required radiation calculations for precipitation are also
currently not implemented in the ALCF, even though this is a planned future
addition.  In order to achieve a fair comparison of observations with model
output, we exclude observed and simulated lidar profiles with precipitation
either manually or using an automated method. It is relatively difficult to
distinguish precipitation backscatter from cloud backscatter in lidar
observations, especially when only one wavelength channel and no polarised
channel are available. In models, the same can be accomplished relatively
easily by excluding profiles exceeding a certain surface precipitation flux. In
the observations, using precipitation flux measurements from rain gauges can be
very unreliable on ships due to ship movement, turbulence caused by nearby ship
structures, and sea spray. Our analysis of rain gauge data from the RV
\emph{Tangaroa} showed large discrepancies between the rain gauge time series
and human-performed synoptic observations, as well as large inconsistencies in
the rain gauge time series. Human-performed observations of precipitation
presence or absence are expected to be reliable but only cover a limited set of
times. Therefore, it was desirable to implement a method of detecting
precipitation from observed backscatter profiles alone.

On the RV \emph{Polarstern} voyages, regular human-performed synoptic
observations were available and included precipitation presence or absence and
type. We used this dataset to train a convolutional artificial neural network
(ANN) to recognise profiles with precipitation from lidar backscatter (Fig.
\ref{fig:ann}a), implemented in the TensorFlow ANN framework
\citep{tensorflow}. Samples of short time intervals (10 min) of near-surface
lidar backscatter (0–250 m) were classified as clear, rain, snow, and fog,
using the synoptic observations as a training dataset (Fig.  \ref{fig:ann}b).
From these, a binary, mutually exclusive classification of profiles as
precipitating (rain or snow) or dry (clear or fog) was derived.  For detecting
model and reanalysis precipitation, we used a fixed threshold for surface
precipitation flux of 0.1 mm h$^{-1}$ (the ANN was not used).

The ANN achieved 65\% sensitivity and 87\% specificity when the true positive
rate (26\%) was made to match observations. The receiver operating
characteristic curve is shown in Fig. \ref{fig:ann}c. We considered these rates
satisfactory for the purpose of filtering precipitation profiles. Fig.
\ref{fig:ann}d shows examples of the predicted precipitation compared to
human-performed observations. The main ANN (`ANN` in Fig. \ref{fig:ann}) was
trained on all data, and ancillary ANNs (`ANN2` in Fig \ref{fig:ann}) were
trained on all data except for each voyage. These were used for testing the
results relative to each voyage.

\begin{figure}[b!]
\centering
\includegraphics[width=\textwidth]{img/lts_dist.pdf}
\caption{
Lower tropospheric stability (LTS) distribution in \textbf{(a)} ERA5 and
\textbf{(b)} MERRA-2 calculated for the 31 voyage tracks and one station from
the highest instantaneous temporal resolution data available. Shown is also the
chosen dividing threshold of 12 K for conditions weak and strong stability.
}
\label{fig:lts}
\end{figure}

\subsection{Partitioning by cyclonic activity and stability}
\label{sec:cyclone-stability}

We partitioned our data into two mutually exclusive subsets by cyclonic
activity. For this purpose, we used a cyclone tracking algorithm to identify
extratropical and polar cyclones (ECs and PCs) over the SO in the reanalysis
and ICON data. We used the open source cyclone tracking package CyTRACK
\citep{perez-alarcon2024}.  Generally, what constitutes an EC is considered
relatively arbitrary due to the very large variability of ECs \citep{neu2013}.
We used the mean sea level pressure field and horizontal wind speed fields as
input to the CyTRACK algorithm. The algorithm uses pressure and wind speed
thresholds as well as tracking across time steps to identify cyclone centres
and radii in each time step. With this information, we could classify
geographical areas as either cyclonic or non-cyclonic. Due to a relatively
small total area covered by cyclones (as identified by the cyclone centre and
radius), we chose a circle of a double radius (relative to one identified by
CyTRACK) centred at the cyclone centre as a cyclonic area for every time step
and cyclone. All other areas were identified as non-cyclonic. For identifying
cyclones in the observations and the reanalyses, ERA5 pressure and wind fields
were used as the input to CyTRACK.  This is justified by the fact that the
large-scale pressure and wind fields in ERA5 are likely sufficiently close to
reality. For identifying cyclones in ICON, its own pressure and wind fields
were used as the input to CyTRACK, because the model is free-running, and thus
the pressure and wind fields are different from reality. Subsetting by
proximity to cyclones is a relatively crude measure because it does not take
into account the different sectors of cyclones, which are commonly associated
with different weather situations. However, this was a choice made for
simplicity of the analysis, given the amount of data.

In addition to the above, we partitioned our data into two mutually exclusive
subsets based on LTS, which is derived as the difference between the potential
temperature at 700 hPa and the surface.  Based on a histogram of LTS in ERA5
and MERRA-2 calculated at all voyage tracks and stations (Fig.  \ref{fig:lts}),
we determined a statistically-based dividing threshold of 12 K for weak
stability (< 12 K) and strong stability (>= 12 K) conditions.

\section{Results}
\label{sec:results}

\begin{figure}[p!]
\centering
\includegraphics[width=\textwidth]{img/cyc_stab_dist.pdf}
\caption{
Geographical distribution of \textbf{(a, b)} cyclonic days and \textbf{(b, d)}
strong stability (LTS > 12 K) time steps in \textbf{(a, c)} ERA5 in years
2010--2013 (inclusive) and \textbf{(b, d)} ICON in model years 2021--2023 (free
running). Cyclonic days are expressed as a fraction of the number of days with
cyclonic activity, defined as grid points located within a double radius of any
cyclone on a given day (UTC), as identified by CyTRACK.
}
\label{fig:cyclone-stability}
\end{figure}

\subsection{Cyclonic activity and stability}

Fig. \ref{fig:cyclone-stability}a, b show a geographical distribution of the
fraction of cyclonic days as determined by the cyclone tracking algorithm
applied to the ERA5 reanalysis and ICON data (Section
\ref{sec:cyclone-stability}). As expected, the strongest cyclonic activity is
in the high-latitude SO zone, and it is relatively zonally symmetric at all
latitudes.  The pattern matches reasonably well with \cite{hoskins2005}. While
both reanalysis and the model agree relatively well, there are clear
differences particularly in the highest occurrence rate regions -- over the
Amundsen Sea, where the Amundsen Sea low is located \cite{turner2013} and which
is more cyclonic in the reanalysis, and around Cape Adare, which is more
cyclonic in ICON. These differences might, however, stem from the relatively
short time periods of comparison (4 years) and the fact that the model is
free-running.

Fig. \ref{fig:cyclone-stability}c, d show a geographical distribution of the
conditions of weak and strong stability as determined by the LTS (Section
\ref{sec:cyclone-stability}). Conditions of weak stability are prevalent in the
mid-to-high SO (with respect to our SO partitioning; 50--65°S), which might be
explained by the relatively cold near-surface air overlying the relatively warm
sea surface. Conditions of strong stability are prevalent elsewhere over the
SO. The distribution is also less zonally symmetric than the cyclonic activity.
In the high-latitude SO, the presence of sea ice might have substantial
stabilising effect \citep{knight2024}. The ERA5 reanalysis is also
substantially more stable than ICON across the whole region.

\begin{figure}[p!]
\centerline{
\includegraphics[width=1.06\textwidth]{img/cloud_occurrence_panel.pdf}
}
\caption{
Cloud occurrence by height for the 31 voyages and one sub-Antarctic
station (MICRE) in observations (O) and simulated by the ALCF from the ICON
model (I), MERRA‐2 (M), and ERA5 reanalysis data (E). The numbers in the legend
indicate the total cloud fraction and the number of days of data. Multiple lines
of ICON profiles are for each of the four years of model data available.
}
\label{fig:cloud-occurrence-panel}
\end{figure}

\subsection{Cloud occurrence by height}
\label{sec:cloud-occurrence}

We used the ALCF to derive cloud occurrence by height and the total cloud
fraction from observations, ICON, ERA5 and MERRA-2 (Fig.
\ref{fig:cloud-occurrence-panel}). In addition, we aggregated the data sources
(voyages and stations) by calculating the averages and percentiles of all
individual profiles, presented in Fig. \ref{fig:cloud-occurrence}. The analysis
shows that the total cloud fraction (determined as the fraction of profiles
with clouds at any height in the lidar cloud mask) is underestimated in ICON by
about 10\% and in the reanalyses by about 20\%. When analysed by height, ICON
overestimates cloud occurrence below 1 km and underestimates it above, MERRA-2
underestimates cloud occurrence at all heights by up to about 10\%, especially
near the surface, and ERA5 simulates cloud occurrence relatively well above 1
km, but strongly underestimates it near the surface.  We note that fog or
near-surface clouds are strongly underestimated in the reanalyses (fog and
clouds are both included in the cloud occurrence).  As shown in Fig.
\ref{fig:cloud-occurrence-panel}, the biases are relatively consistent across
the data sources and longitudes. We conclude that the ICON results  the
observations better than the reanalyses in this metric.

For all observations considered (Fig. \ref{fig:cloud-occurrence}a), the data
show cloud occurrence peaking at the surface, whereas the models show a higher
peak (at about 500 m). The models generally underestimate the total cloud
fraction by 10--20\% and show a strong reduction in cloud occurrence near the
surface, which is not supported by the observations. ICON and ERA5 overestimate
cloud occurrence at their peak (between 0--1 km). Above 1 km, ICON and MERRA-2
underestimate cloud occurrence, but ERA5 is accurate to about 3\% or less. The
exaggerated peak in models is partly supported by the lifting condensation
level (LCL) distribution, which peaks higher by about 200 m in the models than
in the observations (at the surface), although this is not very pronounced.
This is indicative of near-surface relatively humidity being often close to
saturation in the observations but not in the models.

When subsetted by latitude (Fig.  \ref{fig:cloud-occurrence}b, c), we see that
the low-latitude SO zone displays a stronger peak of cloud occurrence near the
surface than the high-latitude SO zone, and this could be because higher
latitudes have less stable atmospheric profiles. The low- and high-latitude SO
zones show similar biases in models as in the general case, but ERA5 does not
overestimate the peak in the low-latitude SO zone (near-surface cloud
occurrence is still strongly underestimated).

When subsetted by cyclonic and non-cyclonic situations (Fig.
\ref{fig:cloud-occurrence}d, e), we see that the cyclonic situations have a
larger amount of observed cloudiness, including peak and total cloud fraction,
both by about 7\%.  In the cyclonic situations, the model vertical profiles of
cloud occurrence compare well with observations, but they peak higher by about
200 m and stronger by about 8\%. The reanalyses still tend to underestimate
cloud occurrence above 1 km by about 5\% and near the surface by about 14\%.
Non-cyclonic situations are similar to the general case, partially also because
they form the majority of cases.

When subsetted by conditions of weak and strong stability (Fig.
\ref{fig:cloud-occurrence}f, g), as defined in Section
\ref{sec:cyclone-stability}, we see that in situations of strong stability
cloud occurrence peaks strongly at the surface in observations, compared to
situations of weak stability, where the peak is more diffuse between 0 and 1
km.  It is work mentioning that conditions of strong stability might be
associated with the formation of advection fog, such as in situation of warm
air advection from the north over colder sea surface, thus inducing fog
formation by cooling of the warm and humid air by the cold surface.  In
situations of strong stability, the models have smaller biases than in weak
stability, with overestimated peak by up to 12\%, underestimated cloud
occurrence above 1 km by up to 5\% and underestimated cloud occurrence near the
surface by about 11\%.  In situations of weak stability, the bias in ICON is
very pronounced, with a much stronger peak at about 500 m, ERA5 underestimates
cloud occurrence below 1 km (especially near the surface), and MERRA-2
underestimates cloud occurrence even more strongly.

In all situations, even when the models overestimate cloud occurrence at some
altitudes, they always substantially underestimate the total cloud fraction.
ICON can be generally characterised as substantially overestimating cloud
occurrence below 1 km and underestimating above, underestimating the total
cloud fraction, and showing greatest biases in conditions of weak stability and
non-cyclonic conditions. It also shows a peak of cloud occurrence at higher
altitude than observations (500 m vs. near the surface), and correspondingly,
its LCL tends to be higher. MERRA-2 can be generally characterised as
underestimating cloud occurrence at nearly all altitudes as well as the total
cloud fraction, but mostly above and below 500 m (the peak at 500 m is
well-represented). MERRA-2 displays the largest errors relative to observations
in the low-latitude SO zone and in situations of weak stability. ERA5 can be
generally characterised as representing cloud occurrence correctly above about
1--1.5 km, overestimating below, but underestimating near-surface cloud
occurrence (0--500 m). The total cloud fraction is strongly underestimated in
all situations. ERA5 has a tendency towards underestimation in the low-latitude
SO zone and situations of weak stability; conversely, overestimating in the
high-latitude SO zone and conditions of strong stability.

\begin{figure}[p!]
\centering
\includegraphics[width=\textwidth]{img/cl_agg.pdf}
\caption{
Cloud occurrence by height calculated as the average of all voyages and
stations and lifting condensation level (LCL) distribution. The LCL is derived
from radiosonde profiles and equivalent model profiles, which were not
available for all voyages and times. The total cloud fraction (CF), average
shortwave (SW), and longwave (LW) and the relative frequency of occurrence
(RFO) are shown. The bands are the 16$^\mathrm{th}$--84$^\mathrm{th}$
percentile calculated from the set of all voyages and stations.
}
\label{fig:cloud-occurrence}
\end{figure}

\subsection{Cloud cover}
\label{sec:cloud-cover}

We also analysed the daily cloud cover (total cloud fraction) distribution.
This is a measure of cloudiness, irrespective of height, calculated over the
course of a day (UTC). A cloud detected at any height means that the lidar
profile was classified as cloudy; otherwise, it was classified as clear sky.
When all profiles in a day are taken together, the cloud cover for the day is
defined as the fraction of cloudy profiles in the total number of profiles,
expressed in oktas (multiples of 1/8).

In Fig.  \ref{fig:cloud-cover} we show the results for the same subsets of data
as in Section \ref{sec:cloud-occurrence}. Observations display the highest
proportion of high cloud cover values (5--8 oktas), peaking at 7 oktas. This
pattern is not represented by ICON or either reanalyses.  While ICON is closest
to matching the observed distribution, it tends to be 1 okta clearer than the
observations, peaking at 6 oktas, and substantially underestimating days with 8
oktas.  Overall, the reanalyses show results similar to each other,
underestimating cloud cover by about 2 oktas and strongly underestimating days
with 7 and 8 oktas. Of the two reanalyses, MERRA-2 has slightly higher cloud
cover then ERA5, by about 6\% at 6 octas, which makes it more consistent with
observations.

When analysed by subsets, observations in the cyclonic subset show the highest
cloud cover, with 8 oktas occurring on one half of such days (Fig.
\ref{fig:cloud-cover}d).  This sensitivity to cyclonic conditions is not
observed in ICON or the reanalyses. Interestingly, clear sky days (0 oktas)
also have a local maximum peaking at about 15\% in this subset.  When we
contrast the low- and high-latitude zones, we see that the high-latitude zone
tends to have greater cloud cover, peaking at 8 oktas (Fig.
\ref{fig:cloud-cover}c). The high-latitude zone also has almost no clear sky or
small cloud cover cases (0--4 oktas). ICON and the reanalyses represent this
characteristic of the distribution well for 0--3 oktas, but otherwise show
biases similar to the general case. One of the greatest biases is present in
ERA5 in the subset of weak stability, in which ERA5 peaks at 3 oktas, while the
observations peak at 7 oktas and show negligible cloud cover below 5 oktas.

\begin{figure}[p!]
\centering
\includegraphics[width=\textwidth]{img/clt_hist.pdf}
\caption{
Daily total cloud fraction histograms calculated as the average of all voyage
and station histograms. The total cloud fraction of a day (UTC) is calculated
as a fraction of cloudy (based on the cloud mask) observed (OBS) or simulated
lidar profiles. The models and subsets are as in Fig.
\ref{fig:cloud-occurrence}.
}
\label{fig:cloud-cover}
\end{figure}

\begin{figure}[p!]
\centering
\includegraphics[width=0.72\textwidth]{img/theta_hur.pdf}
\caption{
Virtual potential temperature ($\theta_v$) and relative humidity (RH)
determined from radiosonde launches and co-located profiles in ICON, ERA5, and
MERRA-2 in subsets as in Fig. \ref{fig:cloud-occurrence}.  The solid lines are
the average calculated from the averages of every individual voyage and
station. The bands span the 16$^\mathrm{th}$--84$^\mathrm{th}$ percentiles
calculated from the distribution of the voyage and station averages. Shown is
also the relative frequency of occurrence and the number of profiles in each
subset.
}
\label{fig:potential-temperature}
\end{figure}

\subsection{Thermodynamic profiles}

In order to link the cloud biases to the local physical conditions, we analysed
about 2300 radiosonde profiles south of 40°S from the 24 RV \emph{Polarstern}
voyages, MARCUS, NBP1704, TAN1702, and TAN1802. Spatially and temporally
colocated profiles were taken from ICON and the reanalyses.  Because the time
period covered by the ICON model output (2021--2024) was different from the
time period covered by the observations (2010--2021), when comparing with the
model, we first had to remap the observation time to model time by taking the
same time relative to the start of the year. Consequently, we also had four
virtual/model profiles (one for each year of 2021--2024) for each observed
profile.  The profiles were partitioned into the same subsets as above
(Sections \ref{sec:cloud-occurrence} and \ref{sec:cloud-cover}).  Apart from
relative humidity, we focus on comparing virtual potential temperature
($\theta_v$) due to its role in low-level tropospheric stability, being one of
the primary factors affecting shallow convection and the associated low-level
cloud formation and dissipation. The observed and model profiles of virtual
potential temperature are shown in Fig. \ref{fig:potential-temperature}.

Overall, the mean $\theta_v$ is accurate less than 0.5 K in ICON and MERRA-2,
being only colder by up to 2 K in the mid-to-high troposphere (less stable) in
ICON than in observations (Fig. \ref{fig:potential-temperature}a).  Larger
differences exist, however, in the 40--55°S zone, where ICON is colder by about
5 K at higher altitudes (Fig.  \ref{fig:potential-temperature}b). In other
subsets, the bias is relatively small. MERRA-2 is very close to the
observations, possibly due to a high accuracy of assimilation of this quantity.
Notably, the variability of virtual potential temperature (as represented by
the percentiles) is much smaller in ICON than in the observations. This
indicates that the model's internal variability in the lower-tropospheric
thermodynamic conditions in the SO is smaller than in reality.

Relative humidity displays much larger biases. In all subsets, ICON is too
humid in the first 1 km by about 5\% but very accurate above, except for the
40--55°S zone and conditions of weak stability (Fig.
\ref{fig:potential-temperature}b, g),  where it is too dry between about 1 and
3 km. MERRA-2, on the other hand, is more humid than observations at all
altitudes and in all subsets, by up to about 20\% at 5 km. Even though the mean
near-surface relative humidity is similar to the observations (Fig.
\ref{potential-temperature}), the distribution in observation is more spread
out across both high and low values, and thus observations have a greater
prevalence of relative humidity close to 100\% and thus LCL located at the
surface (Fig. \ref{cloud-occurrence}a). In our calculations, LCL is a exclusive
function of near-surface temperature, near-surface relative humidity and
surface pressure.

\section{Limitations of this study}

Let us consider the main limitations of the presented results. The spatial
coverage of our dataset does not include most parts of the Indian Ocean and
Pacific Ocean sectors of the SO. Even though climatological features of the SO
are typically relatively uniform zonally, variations exist, such as those
related to the Antarctic Peninsula and the southern tip of South America. The
voyages were mostly undertaken in the Austral summer months and only rarely in
the winter months, due to the poor accessibility of this region during winter.
Therefore, our results are likely representative of summer and to a lesser
extent, spring and autumn conditions.

The time period of ICON is relatively short, with only four full years of
simulation available. Moreover, the simulation is free-running, which means
that observations had to be temporally mapped to this time period (at the same
time relative to the start of the year) for the comparison. For these reasons,
one can expect the results to be slightly different due to reasons unrelated to
model biases, such as different weather conditions, partially accounted for by
the cyclone and stability subsetting, and the phase of climate oscillations
such as the ENSO in the observations and the model.

Ground-based lidar observations are affected by attenuation by thick cloud
layers, and for this reason the results are most representative of boundary
layer clouds, while higher-level clouds are only occasionally visible to the
lidar when boundary layer clouds are not present. Ground-based lidar
observations can be regarded as superior to satellite lidar observations for
low-level clouds, which are predominant in this region, while mid- and
high-level clouds are better sampled by satellite observations
\citep{mcerlich2021}.

We have attempted to remove lidar profiles with precipitation, which could not
be properly simulated with the lidar simulator (Section \ref{sec:ann}).
However, the approach was limited by the relatively low sensitivity of the ANN
(65\%) and the fact that we had to choose a fixed threshold for surface
precipitation flux in the model and reanalyses, which might not correspond to
detection by the ANN applied to observations. Also, we made no attempt to
remove profiles with precipitation not reaching the surface. The above reasons
may result in an artificial bias in the comparison, even though we expect this
to be much smaller than the identified model biases.

\section{Discussion and conclusions}

We analysed a total of about 2400 days of lidar and 2300 radiosonde
observations from 31 voyages/campaigns and a subantarctic station, covering the
Atlantic, Australian, and New Zealand sectors of the SO over the span of 10
years. This dataset, together with the use of a ground-based lidar simulator,
provided a comprehensive basis for evaluating SO cloud and thermodynamic
profile biases in the GSRM ICON and the ERA5 and MERRA-2 reanalyses. Our
analysis provides a unique evaluation perspective different from satellite
observations -- one that we argue is more suitable for evaluating boundary
layer clouds, which are predominant in this region. Furthermore, we subsetted
our dataset by low and high latitude bands, cyclonic activity, and stability in
order to identify how these conditions influence the biases.

Our main finding corroborates previous findings of large boundary layer cloud
biases in models and their subsequent effect on the radiative transfer. This
also applies to the new GSRM ICON, but the biases are generally lower than in
the reanalyses, despite the reanalyses having the advantage of assimilation of
the observed meteorological conditions.  The GSRM has, on the other hand, the
advantage of a much higher spatial resolution and, to a limited extent,
explicit calculation of traditionally subgrid-scale processes such as
convection.

We show that relative to ERA5, the distribution and strength of cyclonic
activity over the SO is well represented in ICON, but it displays lower values
of LTS. The latter is also manifested in the radiosonde profile comparison,
showing that the virtual potential temperature profiles in ICON are less stable
than in the observations over low-latitude SO.

The 31 voyages and a station show remarkably similar biases in cloud occurrence
by height in the lidar comparison, which indicates that common underlying
causes for the biases exist regardless of longitude and season. ICON
underestimates the total cloud fraction by about 10\%, with an overestimation
of clouds below 2 km and an underestimation of clouds above 2 km. The
reanalyses also underestimate the total cloud fraction by about 20\%.  ERA5
overestimates cloud below 1 km but underestimates near-surface cloud or fog.
ICON strongly overestimates the peak of cloud occurrence at about 500 m, which
might be explained by the radiosonde comparison, showing that it is too moist
at around this height.  Similar to our results, \cite{cesana2022} showed that
CMIP6 models also tend to underestimate cloud occurrence above 2 km over the
SO, although their analysis in this case was limited to liquid clouds.

Compared to lidar observations, the daily cloud cover tends to be about 1 okta
lower in ICON and 2 oktas lower in the reanalyses.  Conditions of weak
stability are associated with some of the greatest biases, especially in ERA5.
The models also underestimate the cloud cover very strongly in cyclonic
conditions, which are very cloudy in the observations (8 oktas), but much less
so in the models.

The radiosonde observations indicate that the LCL is too high in ICON and
reanalyses, which is probably responsible for the higher peak of clouds in the
models and the lack of near-surface clouds or fog. The radiosonde comparison,
however, does not seem to explain cloud biases at higher altitudes.  MERRA-2 is
too moist at all heights.  ICON also exhibits smaller internal variability than
the radiosonde observations. Overall, the radiosonde comparison only partially
explains the identified cloud biases, and other physical causes are likely
contributing.  This warrants further investigation, especially of
ocean--atmosphere fluxes, shallow convection, and boundary layer turbulence.
The lack of parametrised subgrid-scale convection in ICON could be a
substantial issue even at the 5-km resolution.

The relationship between cloud biases and radiation has a number of notable
features. Perhaps unsurprisingly, the reanalyses exhibit the too few, too
bright bias previously identified in models. In our results, this is
characterised by outgoing TOA SW radiation similar to or higher than in the
satellite observations, while at the same time total cloud fraction is
substantially underestimated relative to the ground-based lidar observations.
This feature seems to be much more pronounced in ERA5 than in MERRA-2. On the
other hand, this relationship is not present in ICON. This model generally
predicts smaller outgoing TOA SW radiation and smaller total cloud fraction
than observations, and the deficit of outgoing TOA SW radiation is
approximately proportional to the deficit of the total cloud fraction. While
this might be a welcome feature and an improvement over previous models, it
does mean that the outgoing TOA SW radiation is overall underestimated instead
of being compensated by a higher cloud albedo.  This can, of course, lead to
undesirable secondary effects such as overestimated solar heating of the sea
surface, among other factors responsible for SO SST biases in climate models
\citep{zhang2023,luo2023,hyder2018}.  To some extent, the cloud albedo might be
reduced in the model artifically by the application of an inhomogeneity factor
to lower cloud liquid water in the radiative transfer calculations (Sec.
\ref{sec:icon}).

The results imply that SO cloud biases are still a substantial issue even in
the km-scale resolution ICON model, even though an improvement over the
lower-resolution reanalyses is notable. More effort is needed to improve the
model cloud simulations in this understudied region. However, this analysis
suggests that the transition from models with parametrised convection and
clouds to storm-resolving models might not solve these biases without
additional effort.  Evaluation of ocean--atmosphere heat, moisture, and
momentum fluxes against in-situ observations over the SO and comparison of GSRM
simulations against large-eddy simulations are two potential avenues for future
research that could elucidate the physical mechanisms behind the biases, in
addition to the more common efforts in SO cloud microphysics and precipitation
evaluation.

\fontsize{12pt}{14pt}\selectfont
\section*{Acknowledgements}

PK and FB, and the nextGEMS project received funding from the European Union’s
Horizon 2020 research and innovation program under a grant agreement no.
101003470. FB received funding from the Wenner-Gren Foundation and the Swedish
e-Science Research Centre.  The work of GM was supported by the United States
(U.S.) Department of Energy Award DE-SC0021159.  Supercomputing resources were
provided by the DKRZ (project 1125 ICON-development) and the National Academic
Infrastructure for Supercomputing in Sweden (allocation 2023/22-202). The data
collection by the University of Canterbury was funded by the Deep South
National Science Challenge Clouds and Aerosols project.  Data collection on the
AA15-16 voyages was funded by the Australian Antarctic Science project (grant
no.  4292).  We acknowledge the contribution of Thorsten Mauritsen to funding
acquisition and project management.  We acknowledge the RV \emph{Polarstern}
datasets provided by the Alfred Wegener Institute and Pangaea, the AA15-16
dataset provided by the Australian Antarctic Division (AAD) and University of
Canterbury (UC), the RV \emph{Tangaroa} datasets provided by the National
Institute of Water and Atmospheric Research and UC, the NBP1704 dataset
provided by the National Science Foundation, Cooperative Institute for Research
in Environmental Sciences, University of Colorado and UC, the HMNZSW16 dataset
provided by the Royal New Zealand Navy and UC, the MARCUS dataset provided by
ARM and AAD, and the MICRE dataset provided by ARM, the Australian Bureau of
Meteorology, and AAD.  Technical, logistical, and ship support for MARCUS and
MICRE were provided by the AAD through Australian Antarctic Science projects
4292 and 4387, and we thank Steven Whiteside, Lloyd Symonds, Rick van den
Enden, Peter de Vries, Chris Young, Chris Richards, Andrew Klekociuk, John
French, Terry Egan, Nick Cartwright, and Ken Barrett for all of their
assistance.  We thank the scientific staff, the crew, and everyone involved in
collecting data on the voyages and stations, especially Gert König-Langlo,
Holger Schmithüsen, Roger Marchand, Peter Guest, Kelly Schick, Jamie Halla, and
Mike J.  Harvey (†). We thank Loretta Preis for providing additional RV
\emph{Polarstern} data.  We acknowledge the ICON model output provided by the
nextGEMS project, Deutscher Wetterdienst, Max-Planck-Institute for Meteorology,
DKRZ, Karlsruhe Institute of Technology, and Center for Climate Systems
Modeling; reanalysis dataset ERA5 provided by the Copernicus Climate Change
Service; MERRA-2 provided by the Global Modeling and Assimilation Office; CERES
datasets provided by the NASA Langley Atmospheric Science Data Center
Distributed Active Archive Center; and the Natural Earth dataset provided by
naturalearthdata.com.  Last but not least, we acknowledge the use of open
source software Python, Cython \citep{behnel2011}, TensorFlow
\citep{abadi2016}, Devuan GNU+Linux, parallel \citep{tange2011}, NumPy
\citep{harris2020}, SciPy \citep{virtanen2020}, Matplotlib \citep{hunter2007},
cartopy \citep{cartopy}, pyproj, Inkscape, Bash, GNU Fortran, HDF
\citep{folk1999}, and NetCDF \citep{rew1990}.  We dedicate this study to the
memory of Mike J.  Harvey, who very substantially contributed to obtaining the
atmospheric observations on the RV \emph{Tangaroa} voyages used in this study.

\section*{Code availability}

The ALCF, cl2nc, rstool, lidar precipitation detection code, and our data
processing and plotting code are open source and available at
\url{https://alcf.peterkuma.net}, \url{https://github.com/peterkuma/cl2nc},
\url{https://github.com/peterkuma/rstool}, and
\url{https://github.com/peterkuma/alcf-precip},
\url{https://github.com/peterkuma/icon-so-2024}, respectively. CyTRACK is
available at \url{https://github.com/apalarcon/CyTRACK}. The ICON model is
available at \url{https://gitlab.dkrz.de/icon/icon-mpim}.

\section*{Data availability}

The RV \emph{Polarstern} datasets are openly available on Pangaea
(\url{https://pangaea.de}). The MARCUS and MICRE datasets are openly available
from ARM (\url{https://www.arm.gov}). The MERRA-2 data are openly available
from the NASA Goddard Earth Sciences (GES) Data and Information Services Center
(DISC) (\url{https://disc.gsfc.nasa.gov/datasets?project=MERRA-2}).  The ERA5
data are openly available from the Copernicus Climate Data Store (CDS)
(\url{https://cds.climate.copernicus.eu}). The ICON data are available on the
Levante cluster of the DKRZ
(\url{https://www.dkrz.de/en/systems/hpc/hlre-4-levante}) after registration at
\url{https://luv.dkrz.de/register/}. The CERES products are available from the
project website (\url{https://ceres.larc.nasa.gov}) and the NASA Atmospheric
Science Data Centre (\url{https://asdc.larc.nasa.gov/project/CERES}).  The
TAN1802 data are available on Zenodo
(\url{https://doi.org/10.5281/zenodo.4060237}).

% TODO: The other voyage and station data are openly available on Zenodo ().
% Subsetted model and reanalysis data for the voyage tracks and the station and
% data processed with the ALCF are openly available on Zenodo ().

\section*{Author contributions}

The authors have made the following contributions based on the CRediT taxonomy
(\url{https://credit.niso.org}).  PK: conceptualization, data curation, formal
analysis, investigation, methodology, software, writing – original draft; FAMB:
conceptualization, funding acquisition, project administration, supervision,
and writing – review \& editing; AJM, SPA, GMM, JJC, GEP, SH, SP, SG, and AJS:
data curation, investigation, resources, writing – review \& editing.

%\footnotesize
\fontsize{11pt}{13pt}\selectfont
\setlength{\bibsep}{0.0pt}
\bibliography{manuscript}

\end{document}